---
title: "Uganda historic mean t and precip graphs"
output: html_notebook
---


```{r}
library(ggplot2) 
library(tidyr)
library(dplyr)
library(scales)  #for sec axis
library(berryFunctions) # for the last two maps
```

# 1 Read data

The file below contains monthly data for Uganda between 1979-2014. 

```{r}
data <- read.csv("1979-2019_monthly_Uganda.csv", header = TRUE)
# head(data)
# data format:
# lon | lat | year | month | pr | tasmin | tasmax
```

# 2 Prepare data

Get monthly mean t and rainfall across all years. Approach: 
- aggregate precip and tas by month across all available data

```{r}
country_means<-monthlyMeans(data)
```


# 3 Derive graph 

## 3.1 For entire Uganda:
```{r}
map_hist_mean <- 
  ggplot(data=country_means, aes(x=month)) +
  geom_bar(stat="identity", aes(y=pr))+ 
  geom_point(aes(y = tas* 6), color="#008080", size=2) + 
  geom_smooth(aes(y = tas * 6), color="#008080", size=1) +
  scale_x_continuous(breaks = pretty_breaks()) + # to get rid of doubles on the x axis
  scale_y_continuous(name="Precipitation", sec.axis = sec_axis(~ . / 6, name="Temperature"))+
  theme_bw()+
  theme(strip.text.x = element_text(size = 12, face = 'bold', family="DejaVu Sans")) +
  theme(strip.text.y = element_text(size = 12, face = 'bold', family="DejaVu Sans"))
plot(map_hist_mean)
```

## 3.2 For selected points

Point in the north: lat=3.25, lon=33.25
Point in the south: lat=-0.25, lon=31.25

Approach:
- filter precip and tas data for the specific lon+lat values
- aggregate the data by month across all available data

### Prepare data
```{r}
#narrow down the data
north<-filter(data, lat==3.25, lon==33.25)
south<-filter(data, lat== -0.25, lon==31.25)

north_monthly<-monthlyMeans(north)
south_monthly<-monthlyMeans(south)
```

### Derive climate graphs for specific points
```{r}
climateGraph(north_monthly$tas,north_monthly$pr,main = " ",coltemp = "red3", argtext=list(cex=0.7), ylim=c(1,80), lwd=3, compress=FALSE, textprop=0.4)
mtext("Latitude=3.25, Longitude=33.25", col=1, at=6.5, line=3, cex=1.1, font=1)
mtext("1029 m", col=1, at=6.5, line=1, cex=1, font=1)

climateGraph(south_monthly$tas,south_monthly$pr,main = "  ",coltemp = "red3", argtext=list(cex=0.7), ylim=c(10,80), lwd=3, compress=FALSE, textprop=0.4)
mtext("Latitude= -0.25, Longitude=31.25", col=1, at=6.5, line=3, cex=1.1, font=1)
mtext("1318 m", col=1, at=6.5, line=1, cex=1, font=1)

```

