---
title: "Projected MMEM temperature and precipation"
output: html_notebook
author: "Paula Aschenbrenner"
---

Env setup:
```{r}
library(ggplot2) #for mapping with ggplot
library(scales)  #for colour scales
library(RColorBrewer) #for colour scales
library(rgdal) #reading shape files
```

# 1. Load data

```{r}
filename<-("allmodels_allscen_1971-2100_ISIMIP3_monthly.csv")
monthly=read.csv(filename, header = TRUE)
head(monthly)
# data format:
# year | month | lon | lat | pr | tasmin | tasmax | model | scen
```

# 2. Prepare data
Aggregate to create annual values (separate for pr and tas since once sum and once mean)
```{r}
annual_pr<-aggregate(pr ~ lon+lat+year+model+scen, data=monthly, sum, na.rm=TRUE)
annual_tas<-aggregate(cbind(tasmax,tasmin) ~ lon+lat+year+model+scen, data=monthly, mean, na.rm=TRUE)
```

Combine to one data frame and fix naming
```{r}
annual<-cbind(annual_tas,annual_pr$pr)
colnames(annual)[8]<-"pr"
# data format:
#  lon | lat | year |  model | scen | pr | tasmin | tasmax
```

Create column with tas
```{r}
annual$tas<-((annual$tasmax+annual$tasmin)/2)
# data format:
#  lon | lat | year |  model | scen | pr | tasmin | tasmax | tas
```

Create columns for each period of interest
```{r}
period_30<-annual[annual$year<=2040&annual$year>=2021,]
period_50<-annual[annual$year<=2060&annual$year>=2041,]
period_90<-annual[annual$year<=2100&annual$year>=2081,]
period_hist<-annual[annual$year<=2014&annual$year>=1995,]
```

```{r}
# derive mean for indicators across entire period
period_hist<-aggregate(cbind(pr,tasmax,tasmin,tas) ~ lat+lon+model+scen, data=period_hist, mean, na.rm=TRUE)
period_30<-aggregate(cbind(pr,tasmax,tasmin,tas) ~ lat+lon+model+scen, data=period_30, mean, na.rm=TRUE)
period_50<-aggregate(cbind(pr,tasmax,tasmin,tas) ~ lat+lon+model+scen, data=period_50, mean, na.rm=TRUE)
period_90<-aggregate(cbind(pr,tasmax,tasmin,tas) ~ lat+lon+model+scen, data=period_90, mean, na.rm=TRUE)

# add period name to the data
period_hist$period<-"hist"
period_30$period<-"2030"
period_50$period<-"2050"
period_90$period<-"2090"

# historical data will be used for both IPCC scenarios
period_hist_ssp126<-period_hist #rename scenario to later on plot
period_hist_ssp126$scen<-"ssp126"
period_hist_ssp370<-period_hist
period_hist_ssp370$scen<-"ssp370"
```

Add all in one data frame
```{r}
period<-rbind(period_hist_ssp126,period_hist_ssp370,period_30,period_50,period_90)
```

Create multi model ensemble mean 
```{r}
period_mean<-aggregate(cbind(pr,tasmax,tasmin,tas) ~ lat+lon+scen+period, data=period, mean, na.rm=TRUE)
period_mean$model<-"mean"
```


Create multi model ensemble median
```{r}
period_median<-aggregate(cbind(pr,tasmax,tasmin,tas) ~ lat+lon+scen+period, data=period, median, na.rm=TRUE)
period_median$model<-"median"
```

Create files that show difference to historical period
```{r}
period_hist2<-rbind(period_hist_ssp126, period_hist_ssp370)
period_30_dif<-period_30
period_30_dif[,c(5:8)]<-period_30[,c(5:8)]-period_hist2[,c(5:8)]
period_50_dif<-period_50
period_50_dif[,c(5:8)]<-period_50[,c(5:8)]-period_hist2[,c(5:8)]
period_90_dif<-period_90
period_90_dif[,c(5:8)]<-period_90[,c(5:8)]-period_hist2[,c(5:8)]

period_dif<-rbind(period_30_dif,period_50_dif,period_90_dif)
```

Create multi model ensemble mean of dif
```{r}
period_dif_mean<-aggregate(cbind(pr,tasmax,tasmin,tas) ~ lat+lon+scen+period, data=period_dif, mean, na.rm=TRUE)
period_dif_mean$model<-"mean"
```

Create multi model ensemble median of dif
```{r}
period_dif_median<-aggregate(cbind(pr,tasmax,tasmin,tas) ~ lat+lon+scen+period, data=period_dif, median, na.rm=TRUE)
period_dif_median$model<-"median"

#Uganda
Uganda<-readOGR('./shp/UGA_adm1.shp', stringsAsFactors = FALSE) 
```


# 3. Maps of climate in Uganda
##plot mean annual precipitation for multi model mean
```{r}
map_pr_mean <- 
  ggplot() +
  geom_tile(data = period_dif_mean, aes(x = lon, y = lat, fill = pr)) +
  scale_fill_gradientn("Rainfall change, mm", values = c(0, 0.4,0.6, 1),  # values can twist the legend
                       colours=c(brewer.pal(9,'BrBG')),
                       na.value = 'grey') +
  geom_polygon(data=Uganda, aes(x = long, y=lat, group=group), fill = NA, col="black", size = 0.9) +
  geom_hline(yintercept=seq(-2,5,by=2), linetype="dashed", color = "black") + #, width=unit(0.35, "cm")) +
  geom_vline(xintercept=seq(28,35,by=2), linetype="dashed", color = "black") + #, width=unit(0.35, "cm")) +
  theme_bw()+  xlab("Longitude (?E)") + ylab("Latitude (?N)") + 
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text=element_text(size=10), 
        axis.ticks.length = unit(0.2, "cm"),
        legend.title =element_text(size = 13, face = 'bold', family="DejaVu Sans"),
        legend.text=element_text(size=13),
        legend.key.width = unit(0.35, "cm"), #0.6, "cm"),
        legend.key.height = unit(0.7, "cm"), #unit(0.6, "cm"),
        legend.margin = margin(1, 1, 1, 0),
        # legend.spacing = unit, 
        legend.position = "right")  +
  scale_x_continuous(breaks= seq(28,35,by=2) ) +
  scale_y_continuous(breaks= seq(-2,5,by=2) ) + #seq(5,15,by=5), labels = ylabs ) +
  coord_map(xlim = c(29.5,35), ylim = c(-1.5,4.5)) +
  facet_grid(scen ~ period) + #does the magic and plots different maps per defined variable
  theme(strip.text.x = element_text(size = 12, face = 'bold', family="DejaVu Sans")) +
  theme(strip.text.y = element_text(size = 12, face = 'bold', family="DejaVu Sans"))
plot(map_pr_mean)
```


```{r}
map_pr_dif_mean <- ggplot() +
  geom_tile(data = period_dif_mean, aes(x = lon, y = lat, fill = pr)) +
  scale_fill_gradientn("mm",limits=c(-500,500), values = c(0, 0.4,0.6, 1),  # values can twist the legend
                       colours=c(brewer.pal(9,'BrBG')),
                       na.value = 'grey')+
  geom_polygon(data=Uganda, aes(x = long, y=lat, group=group), fill = NA, col="black", size = 0.9) +
  geom_hline(yintercept=seq(-2,5,by=2), linetype="dashed", color = "black") + 
  geom_vline(xintercept=seq(28,35,by=2), linetype="dashed", color = "black") + 
  theme_bw()+ 
  theme(axis.text=element_text(family="DejaVu Sans", size = rel(1.2)),
        axis.title.x=element_text(margin=margin(t=0.4, unit="cm"), size=13,family="DejaVu Sans"),
        axis.title.y=element_text(margin=margin(r=0.4, unit="cm"), size=13, family="DejaVu Sans"),
        axis.ticks.length = unit(0.2, "cm"),
        legend.title =element_text(size = 13, face = 'bold', family="DejaVu Sans", margin=margin(b=0.4, unit="cm")),
        legend.key.width = unit(0.35, "cm"), 
        legend.key.height = unit(0.8, "cm"),
        legend.text = element_text(family="DejaVu Sans", size = rel(1.2)),
        legend.position = "right")  +
  scale_x_continuous(breaks= seq(28,35,by=2), name= "Longitude (E)") +
  scale_y_continuous(breaks= seq(-2,5,by=2), name="Latitude (N)") + 
  coord_map(xlim = c(29.5,35), ylim = c(-1.5,4.5)) +
  facet_grid(scen ~ period) + #does the magic and plots different maps per defined variable
  theme(strip.text.x = element_text(size = 12, face = 'bold', family="DejaVu Sans")) +
  theme(strip.text.y = element_text(size = 12, face = 'bold', family="DejaVu Sans"))
plot(map_pr_dif_mean)
```

```{r}
map_tas_mean <- ggplot() +
  geom_tile(data = period_dif_mean, aes(x = lon, y = lat, fill = tas)) +
  scale_fill_gradientn("°C", limits=c(0, 4), values = c(0, 0.05,0.3,0.6, 0.8, 1),  # values can twist the legend
                       colours=c('white',brewer.pal(9,'YlOrRd'), "#2e004f"),
                       na.value = 'grey')+
  geom_polygon(data=Uganda, aes(x = long, y=lat, group=group), fill = NA, col="black", size = 0.9) +
  geom_hline(yintercept=seq(-2,5,by=2), linetype="dashed", color = "black") + #, width=unit(0.35, "cm")) +
  geom_vline(xintercept=seq(28,35,by=2), linetype="dashed", color = "black") + #, width=unit(0.35, "cm")) +
  theme_bw()+ 
  theme(axis.text=element_text(family="DejaVu Sans", size = rel(1.2)),
        axis.title.x=element_text(margin=margin(t=0.4, unit="cm"), size=13,family="DejaVu Sans"),
        axis.title.y=element_text(margin=margin(r=0.4, unit="cm"), size=13, family="DejaVu Sans"),
        axis.ticks.length = unit(0.2, "cm"),
        legend.title =element_text(size = 13, face = 'bold', family="DejaVu Sans", margin=margin(b=0.4, unit="cm")),
        legend.key.width = unit(0.35, "cm"), 
        legend.key.height = unit(0.8, "cm"),
        legend.text = element_text(family="DejaVu Sans", size = rel(1.2)),
        legend.position = "right")  +
  scale_x_continuous(breaks= seq(28,35,by=2), name= "Longitude (E)") +
  scale_y_continuous(breaks= seq(-2,5,by=2), name="Latitude (N)") +
  coord_map(xlim = c(29.5,35), ylim = c(-1.5,4.5)) +
  facet_grid(scen ~ period) + #does the magic and plots different maps per defined variable
  theme(strip.text.x = element_text(size = 12, face = 'bold', family="DejaVu Sans")) +
  theme(strip.text.y = element_text(size = 12, face = 'bold', family="DejaVu Sans"))
plot(map_tas_mean)
```

