---
title: "Proj hot days and nights"
output: html_notebook
---

The current notebook should be run together with `Helper functions_hot days nights.Rmd`.

```{r}
library(ggplot2) #for mapping with ggplot
library(scales)  #for colour scales
library(RColorBrewer) #for colour scales
library(rgdal) #reading shape files
library(tidyr)
library(dplyr)
```

# 1. Load data
```{r}
# for first run we need the lat+lon points
# either import the csv or generate it using the line below
points<- read_csv("points.csv", col_types = cols(...1 = col_skip()))
#points<-pointsGenesis(data_ssp370_t)
#write.csv(points, "points.csv")

## Read the shape file
Uganda<-readOGR('./shp/UGA_adm1.shp', stringsAsFactors = FALSE) 
```

If all relevant csvs available, skip to Mapping. Below are instructions for first run.
## First run for ssp370 
```{r}
data_ssp370 <- read.csv("alls_ssp370_2014-2100_daily.csv", header = TRUE)
data_ssp370_t<-select(data_ssp370, lat, lon, year, month, tasmax, tasmin, scen, model) #get rid of irrelevant columns
```

# 2. Prepare data

Count number of hot days/nights and record scenario.
Approach: 
- find daily hot units (days>=35/nights?=25)
- aggregate/count them by lon+lat+year+month (summarize)
- aggregate by ...+year (sum)
- aggregate by lon+lat (mean) per period, recording it
- add missing lon+lat coordinates

```{r}
# Hot Days
data_ssp370_hot_days<-hot_days(data_ssp370_t, TRUE)
data_ssp370_hot_days$scen<-"ssp370"

# Hot Nights
data_ssp370_hot_nights<-hot_nights(data_ssp370_t, TRUE)
data_ssp370_hot_nights$scen<-"ssp370"
```


## Now for ssp126
Load data
```{r}
data_ssp126 <- read.csv("alls_ssp126_2014-2100_daily.csv", header = TRUE)
data_ssp126_t<-select(data_ssp126, lat, lon, year, month, tasmax, tasmin, scen, model) #get rid of irrelevant columns
```
Prepare it
```{r}
# Get count of hot days across all unique lat&lon for year/month
data_ssp126_hot_days<-hot_days(data_ssp126_t, TRUE)
data_ssp126_hot_days$scen<-"ssp126"

data_ssp126_hot_nights<-hot_nights(data_ssp126_t, TRUE)
data_ssp126_hot_nights$scen<-"ssp126"
```

To save all relevant files, if needed
```{r}
#write.csv(data_ssp370_hot_days,"proj_hot_days_ssp370.csv")
#write.csv(data_ssp370_hot_nights,"proj_hot_nights_ssp370.csv")

#write.csv(data_ssp126_hot_days,"proj_hot_days_ssp126.csv")
#write.csv(data_ssp126_hot_nights,"proj_hot_nights_ssp126.csv")
```

# 3. Merge files
Combine hot days and nights for both scenarios
```{r}
data_hot_days<-rbind(data_ssp370_hot_days, data_ssp126_hot_days)
data_hot_nights<-rbind(data_ssp370_hot_nights, data_ssp126_hot_nights)

#write.csv(data_hot_days,"proj_hot_days_all.csv")
#write.csv(data_hot_nights,"proj_hot_nights_all.csv")
```


# 4. Map data

Hot days
```{r}
ggplot() +
geom_tile(data = data_hot_days, aes(x = lon, y = lat, fill = count)) +
  scale_fill_gradientn("Hot days/year", values = c(0, 0.003, 0.05,0.3,0.6, 0.8, 0.9, 1),  # values can twist the legend
                       colours=c('white',brewer.pal(9,'YlOrRd'), "#2e004f"),
                       na.value = 'grey')+
  geom_polygon(data=Uganda, aes(x = long, y=lat, group=group), fill = NA, col="black", size = 0.9) +
  geom_hline(yintercept=seq(-2,5,by=2), linetype="dashed", color = "black") + #, width=unit(0.35, "cm")) +
  geom_vline(xintercept=seq(28,35,by=2), linetype="dashed", color = "black") + #, width=unit(0.35, "cm")) +
  theme_bw()+ 
  theme(axis.text=element_text(family="DejaVu Sans", size = rel(1.2)),
        axis.title.x=element_text(margin=margin(t=0.4, unit="cm"), size=13,family="DejaVu Sans"),
        axis.title.y=element_text(margin=margin(r=0.4, unit="cm"), size=13, family="DejaVu Sans"),
        axis.ticks.length = unit(0.2, "cm"),
        legend.title =element_text(size = 13, face = 'bold', family="DejaVu Sans", margin=margin(b=0.4, unit="cm")),
        legend.key.width = unit(0.35, "cm"), 
        legend.key.height = unit(0.8, "cm"),
        legend.text = element_text(family="DejaVu Sans", size = rel(1.2)),
        legend.position = "right")  +
  scale_x_continuous(breaks= seq(28,35,by=2), name= "Longitude (E)") +
  scale_y_continuous(breaks= seq(-2,5,by=2), name="Latitude (N)") +
  coord_map(xlim = c(29.5,35), ylim = c(-1.5,4.5)) +
  facet_grid(scen ~ period) + #does the magic and plots different maps per defined variable
  theme(strip.text.x = element_text(size = 12, face = 'bold', family="DejaVu Sans")) +
  theme(strip.text.y = element_text(size = 12, face = 'bold', family="DejaVu Sans"))

```

Hot nights
```{r}
ggplot() +
geom_tile(data = data_hot_nights, aes(x = lon, y = lat, fill = count)) +
  scale_fill_gradientn("Hot nights/year", values = c(0, 0.0002, 0.005, 0.03, 0.07, 0.1, 0.4, 0.6, 0.8, 1),  # values can twist the legend
                       colours=c('white',brewer.pal(9,'YlOrRd'), "#2e004f"),
                       na.value = 'grey')+
  geom_polygon(data=Uganda, aes(x = long, y=lat, group=group), fill = NA, col="black", size = 0.9) +
  geom_hline(yintercept=seq(-2,5,by=2), linetype="dashed", color = "black") + #, width=unit(0.35, "cm")) +
  geom_vline(xintercept=seq(28,35,by=2), linetype="dashed", color = "black") + #, width=unit(0.35, "cm")) +
  theme_bw()+ 
  theme(axis.text=element_text(family="DejaVu Sans", size = rel(1.2)),
        axis.title.x=element_text(margin=margin(t=0.4, unit="cm"), size=13,family="DejaVu Sans"),
        axis.title.y=element_text(margin=margin(r=0.4, unit="cm"), size=13, family="DejaVu Sans"),
        axis.ticks.length = unit(0.2, "cm"),
        legend.title =element_text(size = 13, face = 'bold', family="DejaVu Sans", margin=margin(b=0.4, unit="cm")),
        legend.key.width = unit(0.35, "cm"), 
        legend.key.height = unit(0.8, "cm"),
        legend.text = element_text(family="DejaVu Sans", size = rel(1.2)),
        legend.position = "right")  +
  scale_x_continuous(breaks= seq(28,35,by=2), name= "Longitude (E)") +
  scale_y_continuous(breaks= seq(-2,5,by=2), name="Latitude (N)") +
  coord_map(xlim = c(29.5,35), ylim = c(-1.5,4.5)) +
  facet_grid(scen ~ period) + #does the magic and plots different maps per defined variable
  theme(strip.text.x = element_text(size = 12, face = 'bold', family="DejaVu Sans")) +
  theme(strip.text.y = element_text(size = 12, face = 'bold', family="DejaVu Sans"))

```

