---
title: "Uganda heavy precipitation proj"
output: html_notebook
---

```{r}
library(ggplot2) #for mapping with ggplot
library(scales)  #for colour scales
library(RColorBrewer) #for colour scales
library(rgdal) #reading shape files
library(tidyr)
library(dplyr)
library(readr)
```


# 1. Load data

```{r}
referer<- read_csv("hist_95 percentile_pr_Uganda.csv", col_types = cols(...1 = col_skip()))
Uganda<-readOGR('./shp/UGA_adm1.shp', stringsAsFactors = FALSE) # preferred as it is being processed better by ggplot
```

# Scenario ssp370

```{r}
pr_all <- read_csv("alls_ssp370_2014-2100_daily.csv")
```

# 2. Prepare data

Approach for each scenario (!):
- filter wet days
- record the periods
- get .95 percentile for each period
- filter and aggregate heavy precip (>= pr_95) for each period
- compare with historic .95 (take the difference)
```{r}
pr_all<-filter(pr_all, pr>0.1)
pr_all<-select(pr_all, lon,lat,year,Date, pr)
# Note: adapt the function strings, if needed
pr_all<-divideAndRecord(pr_all)   # split and store each period as a separate file to speed up the calculations
# last operation saves data in 3 dataset for respective periods
```

For each period, take quantile and find difference to hist. 
For the sake of memory management, each period should be run separately, clearing the memory once saved locally
```{r}
pr_90 <- read_csv("ssp370_pr_90.csv", col_types = cols(...1 = col_skip()))
# Note: adapt the function strings
pr_90<-seever(pr_90)
pr_90<-precipChange(referer, pr_90) # make sure referer exists


pr_30 <- read_csv("ssp370_pr_30.csv", col_types = cols(...1 = col_skip()))
# Note: adapt the function strings
pr_30<-seever(pr_30)
pr_30<-precipChange(referer, pr_30) 

pr_50 <- read_csv("ssp370_pr_50.csv", col_types = cols(...1 = col_skip()))
# Note: adapt the function strings
pr_50<-seever(pr_50)
pr_50<-precipChange(referer, pr_50) 

## Read previous files that were removed for memory concerns
pr_30 <- read_csv("ssp370_pr_30_95_diff.csv", col_types = cols(...1 = col_skip()))
pr_50 <- read_csv("ssp370_pr_50_95_diff.csv", col_types = cols(...1 = col_skip()))
pr_90 <- read_csv("ssp370_pr_90_95_diff.csv", col_types = cols(...1 = col_skip()))

pr_370<-rbind(pr_30, pr_50, pr_90)
write.csv(pr_370, "pr_ssp370_heavy precipitation.csv")
```

# Repeat for the second scenario
Before running, make sure function string are adapted accordingly, otherwise it will write over existing ones!!
Also delete previous variables. 

Load relevant data
```{r}
pr_all <- read_csv("alls_ssp126_2014-2100_daily.csv")
```
Filter relevant data
```{r}
pr_all<-filter(pr_all, pr>0.1)
pr_all<-select(pr_all, lon,lat,year,Date, pr)
# Note: adapt the function strings, if needed
pr_all<-divideAndRecord(pr_all)   # split and store each period as a separate file to speed up the calculations
# last operation saves data in 3 dataset for respective periods
```
Proceed with data preparation
```{r}
# For each period, take quantile and find difference to hist 
pr_30 <- read_csv("ssp126_pr_30.csv", col_types = cols(...1 = col_skip()))
# Note: adapt the function strings
pr_30<-seever(pr_30)
pr_30<-precipChange(referer, pr_30) # make sure referer exists

pr_50 <- read_csv("ssp126_pr_50.csv", col_types = cols(...1 = col_skip()))
# Note: adapt the function strings
pr_50<-seever(pr_50)
pr_50<-precipChange(referer, pr_50) # make sure referer exists

pr_90 <- read_csv("ssp126_pr_90.csv", col_types = cols(...1 = col_skip()))
# Note: adapt the function strings
pr_90<-seever(pr_90)
pr_90<-precipChange(referer, pr_90) # make sure referer exists

## Read previous files that were removed for memory concerns
#pr_30 <- read_csv("ssp126_pr_30_95_diff.csv", col_types = cols(...1 = col_skip()))
#pr_50 <- read_csv("ssp126_pr_50_95_diff.csv", col_types = cols(...1 = col_skip()))
#pr_90 <- read_csv("ssp126_pr_90_95_diff.csv", col_types = cols(...1 = col_skip()))

pr_126<-rbind(pr_30, pr_50, pr_90)
write.csv(pr_126, "pr_ssp126_heavy precipitation.csv")
# delete all variables
```

# 3. Merge data

Load and combine both files.

```{r}
pr_126<- read_csv("pr_ssp126_heavy precipitation.csv", col_types = cols(...1 = col_skip()))
pr_370<- read_csv("pr_ssp370_heavy precipitation.csv", col_types = cols(...1 = col_skip()))

pr_heavy_all_periods<-rbind(pr_126, pr_370)

# save for later
#write.csv(pr_heavy_all_periods, "proj_heavy precipitation_Uganda.csv")
```

# 4. Map data

```{r}
ggplot() +
  geom_tile(data = pr_heavy_all_periods, aes(x = lon, y = lat, fill = diff)) +
  scale_fill_gradientn("mm", values = c(0, 0.2, 0.3, 0.4, 0.6, 0.8, 1),  # values can twist the legend
                       colours=c(brewer.pal(9,'BuPu')),
                       na.value = 'grey')+
  geom_polygon(data=Uganda, aes(x = long, y=lat, group=group), fill = NA, col="black", size = 0.9) + # for the country map
  geom_hline(yintercept=seq(-2,5,by=2), linetype="dashed", color = "black") + # for y-axis
  geom_vline(xintercept=seq(28,35,by=2), linetype="dashed", color = "black") + # for x-axis
  theme(axis.text=element_text(family="DejaVu Sans", size = rel(1.2)),
        axis.title.x=element_text(margin=margin(t=0.4, unit="cm"), size=13,family="DejaVu Sans"),
        axis.title.y=element_text(margin=margin(r=0.4, unit="cm"), size=13, family="DejaVu Sans"),
        axis.ticks.length = unit(0.2, "cm"),
        legend.title =element_text(size = 13, face = 'bold', family="DejaVu Sans", margin=margin(b=0.4, unit="cm")),
        legend.key.width = unit(0.35, "cm"), 
        legend.key.height = unit(0.8, "cm"),
        legend.text = element_text(family="DejaVu Sans", size = rel(1.2)),
        legend.position = "right")  +
  scale_x_continuous(breaks= seq(28,35,by=2), name= "Longitude (E)") +
  scale_y_continuous(breaks= seq(-2,5,by=2), name="Latitude (N)") +
  coord_map(xlim = c(29.5,35), ylim = c(-1.5,4.5)) +
  facet_grid(scen ~ period)
```


